Category-Level Transfer Learning from Knowledge Base to Microblog Stream for Accurate Event Detection

### 2016-11-10 paper/lncsFormat.tex
submitted version

@@

\intextsep=0pt plus 0pt minus 0pt
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=1.0\columnwidth]{img/hood.pdf}
        \caption{The comparison of the time series between the raw word \textit{hood} and the \textit{Military} related word \textit{hood}, computed on the \textit{Edinburgh twitter corpus}\cite{petrovic2012using}. The rise of document frequency on July 28th, 2011 is corresponding to the event mentioned in \url{https://en.wikipedia.org/wiki/Fort_Hood\#2011_attack_plot}.}
        \label{fig:hood}
\end{figure}

@@

In our paper, to balance the performance and the cost of leveraging knowledge base for event detection, we develop a novel category-level transfer learning method, namely \textsc{TransDetector}.
It consists of three phases: extracting of category-level topics in knowledge base, conducting transfer learning to get category-level topics in microblog stream, and detecting events from category-level word time series, as illustrated in Figure \ref{fig:modelDesc}.
We explain the main idea of the three parts in \textsc{TransDetector}, and leave more technique details in Section \ref{sec:TransDetector}.
(1) For the \textbf{extracting} phase, we propose a structural-guided category-level topics extraction method on the knowledge base, . 
It bases on the following facts. 
The knowledge base has the three fold hierarchical structure, consisting of the taxonomy graph (\textit{class} \(\rightarrow\) \textit{subclass}), the category-page bipartite graph (\textit{class} \(\rightarrow\) \textit{instance}), and the page-content map (\textit{instance} \(\rightarrow\) \textit{content}), as illustrated in Figure \ref{fig:NSinitializaton}.
In terms of concept level, the latter part is finer than the former. 
And the last page-content map goes into the detail at the word level. 
By considering these three parts together, we can extract a class or category's highly related words, and restore them in \textit{category-level topics in knowledge base}.
(2) For the \textbf{conducting transfer learning} phase, we propose a novel probabilistic model CTrans-LDA for transferring the knowledge.
CTrans-LDA works in the bayesian transfer learning way like \cite{dai2007transferring}, by utilizing the extracted \textit{category-level topics in knowledge base} as the informative priors to bridge two data domains. 
After running CTrans-LDA, it labels whether a word in a microblog message should link to a category concept in knowledge base.
Applying CTrans-LDA on more microblogs, it gets the \textit{category-level topics in microblog stream}.
(3) For the \textbf{detecting} phase,

Since the recent states have filtered out meaningless words, the detected events on recent states is much more accurate than other methods which don't do the transfer learning.
\textsc{CTrans-LDA}, \textit{history states}, and \textit{recent states} forms up our proposed event detection method \textsc{LTDetector}.
Our experiment on the real dataset \textit{Edinburgh twitter corpus (30 million tweets)} validates the effectiveness of our proposed method \textsc{LTDetector}.

%In transfer learning, \textit{history states} play an important role at speeding up building the linkage between the word in microblog and the high level concept in knowledge base.

@@

To sum up, the contribution of this paper is mainly in four aspects.
(1) We propose a structure-guided category-level topics extraction method, which exploits the knowledge base's hierarchical structure to extract categories' highly correlated topics. 
(2) We develop a probabilistic model CTrans-LDA for category-level transfer learning, which utilizes the word co-occurrences and transfers the knowledge base's category-level topics into microblogs. 
(3) Events are detected accurately on category-level word time series, due to richer semantics and less noise. 
(4) Experiment verifies the quality of category-level topics extracted from knowledge base, and the further study on the benchmark \textit{Edinburgh twitter corpus} validates the effectiveness of our proposed transfer learning method for event detection.

%(1) We reveal that transfer learning from knowledge base to microblog stream can benefit event detection task.
%(2) We propose the probability model \textsc{NSDetector}. 
%The model can combine the history data and the recent data, further determine the arrival new document whether it represents a new event. 
%More than that, \textsc{NSDetector} has the capability to update the knowledge of what happened.
%(3) The experiment on the Edinburgh twitter corpus which contains 30 million tweets show that \textsc{NSDetector} promotes the accuracy rate by 9\% without sacrificing the recall rate.

@@

\section{\textsc{TransDetector}} 
\label{sec:TransDetector}
In this section, we present a novel category-level transfer learning method for event detection, namely \textsc{TransDetector}. 
Illustrated in Figure \ref{fig:modelDesc}, \textsc{TransDetector} consists of three parts: (1) extracting category-level topics in knowledge base, (2) transfer learning on category-level topics in microblog stream, (3) detecting events from category-level word time series.

\textit{Category-level Topics in Knowledge Base} is a set of tuples extracted from knowledge base, which weighs the importance of given words to the specific category, defined in Definition 1 formally. 
Taking the \textit{military} category in Figure \ref{fig:modelDesc} as an example, the history state of \textit{military} contain the words \textit{army}, \textit{military}, and \textit{shootings} etc.

\begin{comment}
Knowledge base often contains much information about what happened. 
But the existing RDF model\cite{klyne2006rdf} on knowledge base lacks an efficient mechanism to transfer the information from knowledge base to text stream.  
With the help of knowledge base's structure (detailed in sec \ref{subsec:hs_initialization}), we provide a  lightweight model to transfer the information in knowledge base to text stream.

\textit{History state} is the proposed data structure to restore the information about what happened. 
Initially, \textit{history state} extracts category-related history data into a set of tuples, which weighs the importance of given words to the specific category, defined in Definition 1 formally. 
Taking the \textit{military} category in Figure \ref{fig:modelDesc} as an example, the history state of \textit{military} contain the words \textit{army}, \textit{military}, and \textit{shootings} etc.

\textit{Recent state} is the proposed data structure to maintain the information about what's happening.
It extracts category-related words from incoming text stream, for each time window.
Still taking the \textit{military} category in  Figure \ref{fig:modelDesc} as an example, based on extracted history states, our proposed probabilistic continuous learning model CTrans-LDA can recognize the words \textit{libyan} and \textit{rebel} in the time window 2011-07-28 in the text stream are related to the category \textit{military}. 
Furthermore the time series analysis on the neighbouring recent states can recognize the set of events' candidate words, and finally detects the events in the text stream. 
In the example of Figure \ref{fig:modelDesc}, the words \textit{libyan} and \textit{rebel} are related to the \textit{military} event ``\#libya Libyan rebels chief is killed (2011-07-28)''.

\end{comment}

\begin{rmkWithoutSpacing}[\textbf{Category-level Topics in Knowledge Base}] 
Given category \(c\), the category level topic in knowledge base is defined by a set of tuples, in which the first element is the word \(w^{(c)}_i\) related to the category \(c\), and the second element is the chi-square score \(chi(c,w^{(c)}_{i})\) under the category \(c\). 
And we denote \(c\)'s category-level topic in knowledge base as \(\bm{h}_c=\{<w^{(c)}_i,chi(c,w^{(c)}_{i})>\}_{i=1,...,N_c}\).
\end{rmkWithoutSpacing}

\begin{rmkWithoutSpacing}[\textbf{Category-level Topics in Microblog Stream}] 
Given category \(c\), the category level topic in microblog stream at time \(t\) is defined by a set of tuples, and denoted as \(\bm{r}_{c,t}=\{<w^{(c)}_i,n(c,t,w^{(c)}_{i})>\}_{i=1,...,N_c}\), in which word \(w^{(c)}_{i}\) is related to the category \(c\), and \(n(c,t,w^{(c)}_{i})\) is its document frequency in the time window \(t\).
\end{rmkWithoutSpacing}

\begin{rmkWithoutSpacing}[\textbf{Category-level Word Time Series}] 
Given category \(c\) and word \(w\), the category level word time series is a list extracted from \(\{\bm{r}_{c,t}\}_{t=1}^{T}\), and denoted as \(\{n(c,t,w^{(c)}_{i})\}_{t=1}^{T}\).
\end{rmkWithoutSpacing}

@@

\begin{rmkWithoutSpacing}[Event Phrase] 
Event phrase \(\mathcal{C}_{c,t,i}\) is the \(i\)-th combination of words which occurred in the set of events' candidate words \(\mathcal{B}_{c,t}\), and represents the \(i\)-th event happened in time \(t\) under the category \(c\).
\end{rmkWithoutSpacing}

\begin{rmkWithoutSpacing}[Event Related Microblogs] Event related microblogs \(\mathcal{D}_{c,t,i}\) are articles relating to the \(i\)-th event in time \(t\) of category \(c\), and correspond to the event phase \(\mathcal{C}_{c,t,i}\).
\end{rmkWithoutSpacing}

Section \ref{subsec:hs_initialization} explains the details of how to initialize history states from knowledge base.
Section \ref{subsec:rs_initialization} shows how to perform transfer learning for recent states.
And section \ref{subsec:detection} interprets the high accuracy event detection based on the processed recent states.

\subsection{History States' Extraction}\label{subsec:hs_initialization}
In this part, we discuss in detail how to initialize the history states from the given knowledge base. 
The knowledge base such as Wikipedia has the structure of classes, subclasses, instances, and the edges between them. 
This structure usually can be represented as triples in RDF graph\cite{klyne2006rdf}, which is adopted to build DBPedia\cite{auer2007dbpedia} and YAGO\cite{suchanek2007yago} from Wikipedia. 
But the reasoning and maintenance of knowledge on the graph is usually expensive\cite{broekstra2003inferencing,bursztyn2015reasoning}. 
To make a trade-off between cost and performance, we use the lightweight data structure \textit{history state} to represent the knowledge about what happened, which is extracted from the knowledge base.
And the knowledge base's threefold structure \(G^{(0)}\), \(G^{(1)}\), \(G^{(2)}\) benefits the extraction of history states.

@@

\textbf{Page-Content Map \(G^{(2)}\)}. 
For a specific Wikipedia dumps version, the edges  \textit{page} \(\rightarrow\)\textit{content} in \(G^{(2)}\) define a one-to-one mapping.
There are a bulk of history information restored in the wiki text content.
In order to extract the key words in wiki page's content, which distinguish it from other pages, we use the chi-square statistics\cite{yang1997comparative}\cite{liu2009imbalanced} to measure the importance of each word for the specific page.

For a given category in \(G^{(0)'}\), chi-square statistics also can evaluate the importance of each word appeared in text contents. 
For example, the chi-square statistic of the term \textit{shooting} under the category \textit{military} is 2888.7 under chi-square test with 1 degree of freedom.
That means the term \textit{shooting} is highly related to the category \textit{military}.  
%To prevent the overclaim of rare word by chi-square statitistcs, the word frequency is introduced as a supplement\cite{liu2009imbalanced}, as shown in line \ref{alg:line10inNormalStatesInit} of Algorithm \ref{alg:normalStatesInit}.
Finally, we can get the category's history state, which are  composed by the category related words and their corresponding importance.
%history states衡量一个词在考虑完整历史信息的情况下，对于一个类别的重要程度，例如在military类别下面；从某种程度上讲，history state从knowledge base中抽取了对于该类别的重要特征，有助于我们将它应用到text stream当中。
\setlength{\textfloatsep}{3pt}% Remove \textfloatsep
\begin{algorithm}[h]
\scriptsize
\caption{History State Initialization from Knowledge Base}
\label{alg:normalStatesInit}

\KwIn{Taxonomy's Graph \(G^{(0)}\), Category-Page Bipartite Graph \(G^{(1)}\), Page-Content Bipartite Graph \(G^{(2)}\), topic related category node \(c\)}
\KwOut{History state \(\bm{h}_c\) on category \(c\)}
\(Pages(c)\leftarrow \varnothing\), \(\bm{h}_c \leftarrow \varnothing\)\\
DAG \(G^{(0)'} \leftarrow\) Remove Cycles of \(G^{(0)}\) by nodes' HITS-PageRank scores. \label{alg:line2inNormalStatesInit}\\
\(SuccessorNodes(c) \leftarrow \) Breadth-first-traverse(\(G^{(0)'},c\))\label{alg:line3inNormalStatesInit}\\
\For{\(node \in SuccessorNodes(c)\)}{
    \(Pages(c) \leftarrow Pages(c) \cup G^{(1)}.neighbours(node)\) \label{alg:line5inNormalStatesInit}\\
}
Word frequency table \(n(c,.) \leftarrow \) do word count on the text contents of \(Pages(c)\) \\
Word frequency table \(n(All,.) \leftarrow \) do word count on the text contents of all pages in \(G^{(2)}\).\\
\For{word \(w\) in WordFrequencyTable(All).keys()}{
    \(chi(c,w) \leftarrow \) \(w\)'s chi-square statistics on \(WordFrequencyTable(c)\) and \(WordFrequencyTable(All)\).\\
    \(h_{c,w} \leftarrow chi(c,w)\) \label{alg:line10inNormalStatesInit}\\
}
\Return{\(\bm{h}_c\)}
\end{algorithm}

Considering the full Wikipedia's contents, history state evaluates the importance of word to the concerned category accurately. 
For example, in \textit{Military}'s history state, the top  words are \textit{army}, \textit{military}, \textit{air}, \textit{command}, \textit{force}, and \textit{regiment}, etc. 
The document which contains these words is related to the category \textit{Military} with high probability. 
We further discuss how to apply history states to the text stream on the following subsection \ref{subsec:rs_initialization}.

\subsection{Recent States' Maintenance}
\label{subsec:rs_initialization}
In this subsection, we describe how the proposed probabilistic model CTrans-LDA utilizes the history states to learn the recent states from text stream.

There are two facts inspiring CTrans-LDA. 
(1) The topics in the document may contain history-state-like topics. 
As an example, the tweet ``Libyan rebel chief gunned down in Benghazi (2011-07-28)'' contains the topic that is similar to the \textit{Military}'s history state.
(2) The history-state-like topics can reuse the information  stored in the history states.
The word \textit{libyan} in the aforementioned example tweet ranks much higher in the \textit{Military} and the \textit{Middle East} history states than the other categories' history states.
After considering the relatedness between the remaining context and the history states, the learned topics of the example tweet include \textit{Military} and \textit{Middle East}.

@@

In the above generative process, the line 2(a) is the key point to distinguish CTrans-LDA from LDA\cite{blei2003latent}, where \(\bm{\tau}_k\) is defined by Equation(\ref{eq:wikiPrior}).
\(K_{KB}\) is the number of pre-defined history-state-like topics, and \(S_k\) is the set of words appeared in the \(k\)-th topic's history state.
As \cite{wallach2008structured} mentioned that the asymmetric prior distribution can significantly improve the quality of topic modelling, \(\bm{\phi_k} \sim Dir(\bm{\beta}+ \bm{\tau_k})\) incorporates \textit{category-level topics in knowledge base} into the asymmetric prior of the word distribution on topic. 
The effect of \(\bm{\tau_k}\) is obvious, e.g., \(\tau_{Military,army}/\tau_{Military,basketball}=203\) leads to that topic \textit{Military} prefers to contain the word \textit{army} other than \textit{basketball}.
The parameter \(\lambda\) controls how much the learned topics are similar to the history state, which can be chosen by cross-validated grid-search.

@@

To solve CTrans-LDA, the gibbs sampling is adopted to determine the hidden variable \(z_{dn}\) and the model parameter \(\phi_k\).
In the initialization phase of Gibbs sampling for CTrans-LDA, the hidden variable \(z_{dn}\) is initialized to topic \(k\) with probability \(\hat{q}_{k|v}\) as Equation (\ref{eq:initProbability}).
For the word \(v\) that belongs to any history state, \(\hat{q}_{k|v}\) is proportional to its importance in the history state \(\tau_{kv}\) as Equation \ref{eq:initProbability}(a). 
For the new word \(v\) in text stream, \(\hat{q}_{k|v}\) is   set uniformly \(1/(K-K_{\bm{KB}})\) on the other topics as Equation \ref{eq:initProbability}(b,c). 
The initialization makes sure that the learned topics are aligned to the pre-defined history states.
\setlength{\abovedisplayskip}{1pt}
\setlength{\belowdisplayskip}{1pt}
\begin{scriptsize} 
\begin{equation}
\label{eq:initProbability}
\begin{aligned}
\hat{q}_{k|v}=
\left\{ \begin{aligned}
\frac{\tau_{kv}}{\sum_{k=1}^{K}\tau_{kv}} &,\sum_{k}\tau_{kv}>0 & (a)\\
0&, \sum_{k}\tau_{kv}=0 \ and \ k \leq K_{\bm{KB}} & (b)\\
1/(K-K_{\bm{KB}})&,\sum_{k}\tau_{kv}=0 \ and \ k > K_{\bm{KB}} & (c)
\end{aligned}\right.
\end{aligned}
\end{equation}
\end{scriptsize}

%采样公式p(z|.)，z有两部分。
The sampling process uses the conditional probability \(p(z_{dn}=k|.)\propto (n^{(d)}_{dk}+\alpha m_k)(n^{(w)}_{kv}+\tau_{kv}+\beta)/(n^{(w)}_{k,.}+\tau_{k,.}+V\beta)\), where \(n^{(d)}_{dk}\) is the number of words in document \(d\) assigned to topic \(k\), and \(n^{(w)}_{kv}\) is the times of word \(v\) assigned to topic \(k\).
We set \(\alpha=0.1\) and \(\beta=0.01\) according to the suggestion by \cite{tang2014understanding}.

After Gibbs sampling on discrete time windows, CTrans-LDA learned all hidden topics of words in microblog stream.
And for a specific word type \(w\) and its history-state-like topic assignment \(c\) in time window \(t\), its document frequency is counted as \(n(c,t,w^{(c)})\), which is the element of category \(c\)'s \(t\)-th recent state.
Event detection on recent states is discussed in the following subsection.

@@

%Text stream evolves more quickly with time than knowledge base.
%For example, recent state of Military contain .
%In other words, recent state is history-state-like topic timestamp.%在时间轴上的横截面。

\section{Experiments}
%\subsection{Data Sets}
In this subsection, we demonstrate the effectiveness of \textit{history states} initialized from knowledge base and \textit{recent states} learned by transfer learning.

>>>>

%================================================
%% make a tight section title
%% Save the class definition of \subparagraph
\let\llncssubparagraph\subparagraph
%% Provide a definition to \subparagraph to keep titlesec happy
\let\subparagraph\paragraph
%% Load titlesec
\usepackage[compact]{titlesec}
%% Revert \subparagraph to the llncs definition
\let\subparagraph\llncssubparagraph
\usepackage[compact]{titlesec}% shrink all the whitespace around section headings
%================================================

@@

\intextsep=1pt plus 1pt minus 1pt
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=1.0\columnwidth]{img/hood.pdf}
        \caption{The comparison of the time series between the raw word \textit{hood} and the \textit{Military} related word \textit{hood}, computed on the \textit{Edinburgh twitter corpus}\cite{petrovic2012using}. The rise of document frequency on July 28th, 2011 is corresponding to the event mentioned in \url{https://en.wikipedia.org/wiki/Fort_Hood\#2011_attack_plot}.}
        \label{fig:hood}
\end{figure}

@@

In our paper, to balance the performance and the cost of leveraging knowledge base for event detection, we develop a novel category-level transfer learning method, namely \textsc{TransDetector}.
It consists of three phases: extracting category-level topics in knowledge base, conducting transfer learning to get category-level topics in microblog stream, and detecting events from category-level word time series, as illustrated in Figure \ref{fig:modelDesc}.
We explain the main idea of the three parts in \textsc{TransDetector}, and leave more technique details in Section \ref{sec:TransDetector}.
(1) In the \textbf{extracting} phase, we propose a structural-guided category-level topics extraction method on the knowledge base, . 
It bases on the following facts. 
The knowledge base has the three fold hierarchical structure, consisting of the taxonomy graph (\textit{class} \(\rightarrow\) \textit{subclass}), the category-page bipartite graph (\textit{class} \(\rightarrow\) \textit{instance}), and the page-content map (\textit{instance} \(\rightarrow\) \textit{content}), as illustrated in Figure \ref{fig:NSinitializaton}.
In terms of concept level, the latter part is finer than the former. 
And the last page-content map goes into the detail at the word level. 
By considering these three parts together, we can extract a class or category's highly related words, and restore them in \textit{category-level topics in knowledge base}.
(2) In the \textbf{conducting transfer learning} phase, we propose a novel probabilistic model CTrans-LDA for transferring the knowledge.
CTrans-LDA works in the bayesian transfer learning way like \cite{dai2007transferring}, by utilizing the extracted \textit{category-level topics in knowledge base} as the informative priors to bridge two data domains. 
CTrans-LDA labels whether a word in a microblog message should link to a category in knowledge base, or just label it as no-category-related word.
Applying CTrans-LDA on more microblogs, it gets the \textit{category-level topics in microblog stream} and \textit{category-level word time series}.
(3) In the \textbf{detecting} phase, since the words in microblogs have been enriched semantically and the meaningless words are labeled as no-category-related, the event detection on \textit{category-level word time series} is much more accurate than other methods which don't conduct the transfer learning.
Without loss of generality, we use Wikipedia as a representative of the knowledge base in the remaining of this paper.
And the experiment on the \textit{Edinburgh twitter corpus} demonstrates the effectiveness of our proposed \textsc{TransDetector}.

@@

To sum up, the contribution of this paper is mainly in four aspects.
(1) We propose a structure-guided category-level topics extraction method, exploiting the knowledge base's hierarchical structure to extract categories' highly correlated topics. 
(2) We develop a probabilistic model CTrans-LDA for category-level transfer learning, which utilizes the word co-occurrences and transfers the knowledge base's category-level topics into microblogs. 
(3) Events are detected accurately on category-level word time series, due to richer semantics and less noise. 
(4) Experiment verifies the quality of category-level topics extracted from knowledge base, and the further study on the benchmark \textit{Edinburgh twitter corpus} validates the effectiveness of our proposed transfer learning method for event detection.

@@

\section{\textsc{TransDetector}} 
\label{sec:TransDetector}
In this section, we present a novel category-level transfer learning method for event detection, namely \textsc{TransDetector}. 
Illustrated in Figure \ref{fig:modelDesc}, \textsc{TransDetector} consists of three parts: (1) extracting category-level topics in knowledge base, (2) conducting transfer learning to get category-level topics in microblog stream, and (3) detecting events from category-level word time series.
The following definitions are used by our method \textsc{TransDetector}. 
Definition 1 is used in the \textbf{extracting} phase, Definitions 2 and 3 are used in the \textbf{conducting transfer learning} phase, and Definitions 4 to 6 are used in the \textbf{detecting} phase.
Because Definitions 2 to 6 are very easy to understand, what needs to be specified are Definition 1.
\textit{Category-level topics in knowledge base} is a set of tuples extracted from knowledge base, which weighs the importance of given words to the specific category, defined in Definition 1 formally. 
Taking the \textit{Military} category in Figure \ref{fig:NSinitializaton} as an example, the category-level topic of \textit{military} contain the words \textit{army}, \textit{military}, and \textit{shootings} etc, which are highly related to the category \textit{Military}.

\begin{rmkWithoutSpacing}[\textbf{Category-Level Topics in Knowledge Base}] 
Given category \(c\), the category level topic in knowledge base is defined by a set of tuples, in which the first element is the word \(w^{(c)}_i\) related to the category \(c\), and the second element is the chi-square score \(chi(c,w^{(c)}_{i})\) under the category \(c\). 
And we denote \(c\)'s category-level topic in knowledge base as \(\bm{h}_c=\{<w^{(c)}_i,chi(c,w^{(c)}_{i})>\}_{i=1}^{N_c}\).
\end{rmkWithoutSpacing}

\begin{rmkWithoutSpacing}[\textbf{Category-Level Topics in Microblog Stream}] 
Given category \(c\), the category level topic in microblog stream at time \(t\) is defined by a set of tuples, and denoted as \(\bm{r}_{c,t}=\{<w^{(c)}_i,n(c,t,w^{(c)}_{i})>\}_{i=1}^{N_c}\), in which word \(w^{(c)}_{i}\) is related to the category \(c\), and \(n(c,t,w^{(c)}_{i})\) is its document frequency in the time window \(t\).
\end{rmkWithoutSpacing}

\begin{rmkWithoutSpacing}[\textbf{Category-Level Word Time Series}] 
Given category \(c\) and word \(w\), the category level word time series is a list of document frequencies extracted from \(\{\bm{r}_{c,t}\}_{t=1}^{T}\), and denoted as \(\{n(c,t,w^{(c)}_{i})\}_{t=1}^{T}\).
\end{rmkWithoutSpacing}

@@

\begin{rmkWithoutSpacing}[\textbf{Event Phrase}] 
Event phrase \(\mathcal{C}_{c,t,i}\) is the \(i\)-th combination of words which occurred in the set of events' candidate words \(\mathcal{B}_{c,t}\), and represents the \(i\)-th event happened in time \(t\) under the category \(c\).
\end{rmkWithoutSpacing}

\begin{rmkWithoutSpacing}[\textbf{Event Related Microblogs}] Event related microblogs \(\mathcal{D}_{c,t,i}\) are articles relating to the \(i\)-th event in time \(t\) of category \(c\), and correspond to the event phase \(\mathcal{C}_{c,t,i}\).
\end{rmkWithoutSpacing}

Section \ref{subsec:hs_initialization} explains the details of how to extract \textit{category-level topics in knowledge base} by using the three fold hierarchical structure.
Section \ref{subsec:rs_initialization} shows how to conduct transfer learning to get \textit{category-level word time series} in microblog stream.
And section \ref{subsec:detection} interprets the high accuracy event detection based on the processed \textit{category-level word time series}.

\subsection{Extracting Category-Level Topics in Knowledge Base}\label{subsec:hs_initialization}
In this part, we discuss in detail how to extract the \textit{category-level topics in knowledge base} on the given categories. 
The knowledge base such as Wikipedia has the structure of classes, subclasses, instances, and the edges between them. 
This structure usually can be represented as triples in RDF graph\cite{klyne2006rdf}, which is adopted to build DBPedia\cite{auer2007dbpedia} and YAGO\cite{suchanek2007yago} from Wikipedia. 
But the reasoning and maintenance of knowledge on the graph is usually expensive\cite{bursztyn2015reasoning}. 
To make a trade-off between cost and performance, we use the lightweight data structure \textit{topics} to represent the knowledge at category(class) level.
And the knowledge base's threefold structure \(G^{(0)}\), \(G^{(1)}\), \(G^{(2)}\) benefits the extraction of \textit{category-level topics}.

@@

\textbf{Page-Content Map \(G^{(2)}\)}. 
For a specific Wikipedia dumps version, the edges  \textit{page} \(\rightarrow\)\textit{content} in \(G^{(2)}\) define a one-to-one mapping.
There are a bulk of information restored in the wiki text content.
In order to extract the key words in wiki page's content, which distinguish it from other pages, we use the chi-square statistics\cite{yang1997comparative}\cite{liu2009imbalanced} to measure the importance of each word for the specific page.

For a given category in \(G^{(0)'}\), chi-square statistics also can evaluate the importance of each word appeared in text contents. 
For example, the chi-square statistic of the term \textit{shooting} under the category \textit{military} is 2888.7 under chi-square test with 1 degree of freedom.
That means the term \textit{shooting} is highly related to the category \textit{military}.  
%To prevent the overclaim of rare word by chi-square statitistcs, the word frequency is introduced as a supplement\cite{liu2009imbalanced}, as shown in line \ref{alg:line10inNormalStatesInit} of Algorithm \ref{alg:normalStatesInit}.
Finally, we can get the \textit{category-level topic}, which are  composed by the category related words and their corresponding importance.
\setlength{\textfloatsep}{3pt}% Remove \textfloatsep
\begin{algorithm}[h]
\scriptsize
\caption{Extraction of Category-Level Topics in Knowledge Base}
\label{alg:normalStatesInit}

\KwIn{Taxonomy's Graph \(G^{(0)}\), Category-Page Bipartite Graph \(G^{(1)}\), Page-Content Bipartite Graph \(G^{(2)}\), topic related category node \(c\)}
\KwOut{\(c\)'s category-level topic in knowledge base \(\bm{h}_c\)}
\(Pages(c)\leftarrow \varnothing\), \(\bm{h}_c \leftarrow \varnothing\)\\
DAG \(G^{(0)'} \leftarrow\) Remove Cycles of \(G^{(0)}\) by nodes' HITS-PageRank scores. \label{alg:line2inNormalStatesInit}\\
\(SuccessorNodes(c) \leftarrow \) Breadth-first-traverse(\(G^{(0)'},c\))\label{alg:line3inNormalStatesInit}\\
\For{\(node \in SuccessorNodes(c)\)}{
    \(Pages(c) \leftarrow Pages(c) \cup G^{(1)}.neighbours(node)\) \label{alg:line5inNormalStatesInit}\\
}
Word frequency table \(n(c,.) \leftarrow \) do word count on the text contents of \(Pages(c)\) \\
Word frequency table \(n(All,.) \leftarrow \) do word count on the text contents of all pages in \(G^{(2)}\).\\
\For{word \(w\) in WordFrequencyTable(All).keys()}{
    \(chi(c,w) \leftarrow \) \(w\)'s chi-square statistics on \(WordFrequencyTable(c)\) and \(WordFrequencyTable(All)\).\\
    \(h_{c,w} \leftarrow chi(c,w)\) \label{alg:line10inNormalStatesInit}\\
}
\Return{\(\bm{h}_c\)}
\end{algorithm}

Considering the full Wikipedia's contents, the word's score in the \textit{category-level topic} evaluates the importance of word to the concerned category accurately. 
For example, in \textit{Military}'s \textit{category-level topic}, the top  words are \textit{army}, \textit{military}, \textit{air}, \textit{command}, \textit{force}, and \textit{regiment}, etc. 
The document which contains these words is related to the category \textit{Military} with high probability.

\subsection{Conducting Transfer Learning}
\label{subsec:rs_initialization}
In this subsection, we describe how the proposed probabilistic model CTrans-LDA transfers the category-level topics into microblogs.

There are two facts inspiring CTrans-LDA. 
(1) The topics in the document may contain category-level topics. 
As an example, the tweet ``Libyan rebel chief gunned down in Benghazi (2011-07-28)'' contains the topic that is similar to the \textit{Military}'s \textit{category-level topics in knowledge base}.
(2) The \textit{category-level topics in microblog stream} can reuse the information stored in the corresponding \textit{category-level topics in knowledge base}.
The word \textit{libyan} in the aforementioned example tweet ranks much higher in the \textit{Military} and the \textit{Middle East} category-level topics than the other categories' topics.
After considering the relatedness between the remaining context and the \textit{category-level topics in knowledge base}, the learned topics of the example tweet include \textit{Military} and \textit{Middle East}.

@@

In the above generative process, the line 2(a) is the key point to distinguish CTrans-LDA from LDA\cite{blei2003latent}, where \(\bm{\tau}_k\) is defined by Equation(\ref{eq:wikiPrior}).
\(K_{KB}\) is the number of pre-defined categories used for transferring, and \(S_k\) is the set of words appeared in the \(k\)-th category-level topic.
As \cite{wallach2008structured} mentioned that the asymmetric prior distribution can significantly improve the quality of topic modelling, \(\bm{\phi_k} \sim Dir(\bm{\beta}+ \bm{\tau_k})\) incorporates \textit{category-level topics in knowledge base} into the asymmetric prior of the word distribution on topic. 
The effect of \(\bm{\tau_k}\) is obvious, e.g., \(\tau_{Military,army}/\tau_{Military,basketball}=203\) leads to that topic \textit{Military} prefers to contain the word \textit{army} other than \textit{basketball}.
The parameter \(\lambda\) controls how much the knowledge is used for transferring, which can be chosen by cross-validated grid-search.

@@

To solve CTrans-LDA, the gibbs sampling is adopted to determine the hidden variable \(z_{dn}\) and the model parameter \(\phi_k\).
In the initialization phase of Gibbs sampling for CTrans-LDA, the hidden variable \(z_{dn}\) is initialized to topic \(k\) with probability \(\hat{q}_{k|v}\) as Equation (\ref{eq:initProbability}).
For the word \(v\) that belongs to any category-level topic, \(\hat{q}_{k|v}\) is proportional to its importance in the category-level topic \(\tau_{kv}\) as Equation \ref{eq:initProbability}(a). 
For the new word \(v\) in text stream, \(\hat{q}_{k|v}\) is   set uniformly \(1/(K-K_{\bm{KB}})\) on the other topics as Equation \ref{eq:initProbability}(b,c). 
The initialization makes sure that the learned topics are aligned to the pre-defined category-level topic.
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\begin{scriptsize} 
\begin{equation}
\label{eq:initProbability}
\begin{aligned}
\hat{q}_{k|v}=
\left\{ \begin{aligned}
\frac{\tau_{kv}}{\sum_{k=1}^{K}\tau_{kv}} &,\sum_{k}\tau_{kv}>0 & (a)\\
0&, \sum_{k}\tau_{kv}=0 \ and \ k \leq K_{\bm{KB}} & (b)\\
1/(K-K_{\bm{KB}})&,\sum_{k}\tau_{kv}=0 \ and \ k > K_{\bm{KB}} & (c)
\end{aligned}\right.
\end{aligned}
\end{equation}
\end{scriptsize}

%采样公式p(z|.)，z有两部分。
The sampling process uses the conditional probability \(p(z_{dn}=k|.)\propto (n^{(d)}_{dk}+\alpha m_k)(n^{(w)}_{kv}+\tau_{kv}+\beta)/(n^{(w)}_{k,.}+\tau_{k,.}+V\beta)\), where \(n^{(d)}_{dk}\) is the number of words in document \(d\) assigned to topic \(k\), and \(n^{(w)}_{kv}\) is the times of word \(v\) assigned to topic \(k\).
We set \(\alpha=0.1\) and \(\beta=0.005\) according to the suggestion by \cite{tang2014understanding}.

After Gibbs sampling on discrete time windows, CTrans-LDA learned all hidden topics of words in microblog stream.
And for a specific word type \(w\) and its category assignment \(c\) in time window \(t\), its document frequency is counted as \(n(c,t,w^{(c)})\), which is the element of category-level word time series.
Event detection on category-level word time series is discussed in the following subsection.

@@

%Text stream evolves more quickly with time than knowledge base.
%For example, recent state of Military contain .

\section{Experiments}
%\subsection{Data Sets}
In this subsection, we demonstrate the effectiveness of \textsc{TransDetector}, by evaluating the \textit{category-level topics in knowledge base}, \textit{category-level topics in microblog stream} learned by transfer learning, and the effect of event detection.

\textbf{Knowledge Base.} 
We construct the taxonomy graph \(G^{(0)}\), the category-page bipartite graph \(G^{(1)}\) from the latest dump of category links\footnote{\url{https://dumps.wikimedia.org/enwiki/latest/enwiki-latest-categorylinks.sql.gz}} and the page-content map \(G^{(2)}\) from Wikipedia pages\footnote{\url{https://dumps.wikimedia.org/enwiki/latest/enwiki-latest-pages-articles.xml.bz2 }}.
We set \(K_{KB}=50\), which means 50 categories are selected manually to cover the topics of Wikipedia and the target corpus as widely as possible. 
There are two kinds of categories are considered. 
The mid-high categories in the taxonomy graph \(G^{(0)'}\), which are representative, are likely to be selected, such as \textit{Aviation}, \textit{Military}, and \textit{Middle East}, etc.
And the mid categories, which reflect the main interests of the target corpus, are also taken into consideration, such as \textit{American Football}, \textit{Basketball}, and \textit{Baseball}, etc. 
%Other knowledge bases which have the above mentioned threefold structure are also available for conducting.
